---
title: "Movies_data_analysis: 2016-19"
author: "Godfrey"
date: "2023-09-16"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

**Setting Working Directory**
```{r Working_directory}
setwd("C:/Users/CeX/Desktop/R/web_scraping")
list.files()
rm(list=ls()) #clearing envionment
```

**Installing/Loading Packages**

```{r Packages, echo=FALSE}
library(readxl)
library(rvest)
library(httr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggridges)
library(gridExtra)
library(grid)
library(patchwork)
library(wesanderson)
theme_set(theme_pubr())
```

## 1. Data Aggregation
```{r Aggregating_data}
imdb_df16 <- read_csv("2016_popular_movies_data.csv") #2016 data only
imdb_df17 <- read_csv("2017_popular_movies_data.csv") #2017 data only
imdb_df18 <- read_csv("2018_popular_movies_data.csv") #2018 data only
imdb_df19 <- read_csv("2019_popular_movies_data.csv") #2019 data only

all.data <- rbind(imdb_df16,imdb_df17,imdb_df18,imdb_df19)
str(all.data)

#removing col 1 and putting year at the end of the dataset
all.data <- all.data%>%
  mutate(...1 = NULL, Year = as.factor(Year))%>%
  relocate(Year, .after = PG)%>%
  arrange(Year)
head(all.data,5)
```

## 2. Analysis
  2.1. Visualisation

```{r data_analysis}
qplot(data = all.data,Runtime,fill = Genre,bins = 30)+
labs(title = "Movie Runtime Across Genre",
     x = "Runtime (Min)",
     y = "Count")+
  theme_bw()

ggplot(all.data,aes(x=Runtime,y=Rating))+
  geom_point(aes(size=Votes,col=Genre))+
  labs(title = "Movie Rating and Runtime Across various Genre", 
       x = "Runtime (Min)", 
       y = "Movie Rating")+
  theme(panel.background = element_rect(fill="white"), panel.border = element_rect(color = "Black", fill = NA, size = 0.5))

ggplot(all.data,aes(x=Runtime,y=Gross_earnings))+
  geom_point(aes(size=Metascore,colour=Year))+
  labs(title = "Film Gross Earnings Vs Runtime Across Genre",
     x = "Runtime (Min)", 
     y = "Gross Earnings (USD$ mn)")+
  scale_fill_manual(values=wes_palette(n= 4, name="GrandBudapest1", type = "continuous"))

ggplot(all.data, aes(x=Year, y=Runtime)) +
  geom_jitter(aes(color = Genre, shape = Genre, width = 0.01, size = Runtime))+
  theme_minimal()+
  labs(title = "Runtime distribution across years")

ggplot(all.data,aes(x=Rating,y=Votes/1000))+
geom_point(aes(size=Gross_earnings,col=Genre))+
labs(title = "Movie Votes Vs Rating Across Genre", 
     x = "Rating", 
     y = "Votes ('000)")+
  theme_minimal()+
  #theme(panel.background = element_rect(fill="white"),
        #panel.border = element_rect(color = "Black",
                                    #fill = NA,
                                   # size = 0.5))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(plot.title = element_text(vjust = 2)) 
```
2.2 Exploring further the relationship between Rating and Votes

```{r Rating-Votes}
#Method - use ggridges
  #changing rating into factor
#plot
ggplot(mutate(all.data,Rating.= as.factor(round(Rating, 0))),
       aes(x = Gross_earnings, y = Rating., fill = stat(x))) +
  geom_density_ridges_gradient(scale = 1.5, rel_min_height = 0.01, gradient_lwd = 2.) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = expand_scale(mult = c(0.01, 0.25))) +
  scale_fill_viridis_c(name = "Revenue", option = "C") +
  labs(
    title = 'Gross Revenues of Movies',
    subtitle = 'Gross Earnings of Popular Movies by Rating: 2016-2019'
  ) +
  theme_ridges(font_size = 10, grid = FALSE) + 
  theme(axis.title.y = element_blank())+
  theme_bw()
```


```{r ggridges2}
ggplot(all.data, aes(x = Runtime, y = Genre, fill = stat(x))) +
  geom_density_ridges_gradient(scale = 1.2, rel_min_height = 0.01, gradient_lwd = 1.) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = expand_scale(mult = c(0.01, 0.25))) +
  scale_fill_viridis_c(name = "Runtime", option = "T") +
  labs(
    title = 'Total Runtime of Movies',
    subtitle = 'Runtime (Min) of popular 2016-19 movies by Genre'
  ) +
  theme_ridges(font_size = 13, grid = TRUE) + 
  theme(axis.title.y = element_blank())

```

```{r}
ggplot(
  all.data, 
  aes(x = Gross_earnings, y = PG)
  ) +
  geom_density_ridges_gradient(
    aes(fill = ..x..), scale = 1.5, rel_min_height = 0.01, gradient_lwd = 1.
    ) +
  scale_fill_gradientn(
    colours = c("blue", "red"),
    name = "Revenue",
    )+
  labs(title = 'Gross Revenues of Movies',
    subtitle = 'Gross revenues (USD) of popular 2016-19 Movies by Genre')+
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = expand_scale(mult = c(0.01, 0.25)))+
  theme_ridges(font_size = 13, grid = TRUE) + 
  theme(axis.title.y = element_blank())

### BARPLOT == creating bins
bp <- na.omit(all.data) %>%
  group_by(Genre) %>%
  group_modify(~ ggplot2:::compute_density(.x$Gross_earnings, NULL)) %>%
  rename(Gross_earnings = x)

ggplot(na.omit(all.data), aes( x = Gross_earnings, y = PG, fill = stat(x))) +
  geom_density_ridges_gradient(stat = "binline", bins = 20, scale = 2, draw_baseline = FALSE) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = expand_scale(mult = c(0.01, 0.25))) +
  scale_fill_viridis_c(name = "Revenue", option = "turbo") +#options include: letters from A, B..., "magma", "viridis","plasma""inferno","cividis","mako","rocket","turbo"
  labs(
    title = 'Gross Revenues of Movies',
    subtitle = 'Gross revenues (USD) of popular 2016-19 Movies by Genre'
  ) +
  theme_ridges(font_size = 13, grid = TRUE) + 
  theme(axis.title.y = element_blank())

ggplot(all.data, aes(x = Runtime, y = Genre, color = Year, point_color = Year, fill = Year)) +
  geom_density_ridges(
    jittered_points = TRUE, scale = .95, rel_min_height = .01,
    point_shape = "|", point_size = 3, size = 0.25,
    position = position_points_jitter(height = 0)
  ) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0), name = "Runtime(min)") +
  scale_fill_manual(values = c("#D55E0050", "#0072B250", "#21779e", "#03a1fc"), labels = c("2016", "2017", "2018","2019")) +
  scale_color_manual(values = c("#D55E00", "#0072B2", "#21779e","#03a1fc"), guide = "none") +
  scale_discrete_manual("point_color", values = c("#D55E00", "#0072B2","#21779e","#03a1fc"), guide = "none") +
  coord_cartesian(clip = "off") +
  guides(fill = guide_legend(
    override.aes = list(
      fill = c("#D55E00A0", "#0072B2A0","#21779e","#03a1fc"),
      color = NA, point_color = NA)
  )
  ) +
  ggtitle("Runtime (mn) of Top Popular Movies") +
  theme_ridges(center = TRUE)

```

```{r}
#getting legend
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

#creating plots
y1 <- ggbarplot(subset(all.data,Year==2016), x = "Rank", y = "Gross_earnings",
          fill = "Genre",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "asc",# Sort the value in dscending order
          sort.by.groups = TRUE,     # Don't sort inside each group
          legend = "right",
          x.text.angle = 90,          # Rotate vertically x axis texts,
          ggtheme = theme_pubclean())+
  font("x.text", size = 8, vjust = 0.5)+
  scale_x_discrete(breaks = seq(0,100,5))+labs(x = "2016", y = "Revenue (mn)" )+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10))+
  theme(axis.text.x = element_blank(),  # Remove x-axis labels
        axis.ticks.x = element_blank())  # Remove x-axis tick marks # Remove x-axis labels

y1 <- ggpar(y1, legend.title = "Genre", font.legend = c(8, "plain", "black"))

y2 <- ggbarplot(subset(all.data,Year==2017), x = "Rank", y = "Gross_earnings",
          fill = "Genre",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "asc",# Sort the value in dscending order
          sort.by.groups = TRUE,     # Don't sort inside each group
          x.text.angle = 90,          # Rotate vertically x axis texts
          legend = "none",
          ggtheme = theme_pubclean())+
  font("x.text", size = 8, vjust = 0.5)+
  scale_x_discrete(breaks = seq(0,100,5))+labs(x = "2017", y = "")+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10))+
  theme(axis.text.x = element_blank(),  # Remove x-axis labels
        axis.ticks.x = element_blank())  # Remove x-axis tick marks  # Remove x-axis labels

y3 <- ggbarplot(subset(all.data,Year==2018), x = "Rank", y = "Gross_earnings",
          fill = "Genre",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "asc",# Sort the value in dscending order
          sort.by.groups = TRUE,     # Don't sort inside each group
          x.text.angle = 90,           # Rotate vertically x axis texts
          legend = "none",
          ggtheme = theme_pubclean())+
  font("x.text", size = 8, vjust = 0.5)+
  scale_x_discrete(breaks = seq(0,100,5))+labs(x = "2018", y = "Revenue (mn)")+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10))+
  theme(axis.text.x = element_blank(),  # Remove x-axis labels
        axis.ticks.x = element_blank())  # Remove x-axis tick marks  # Remove x-axis labels

y4 <- ggbarplot(subset(all.data,Year==2019), x = "Rank", y = "Gross_earnings",
          fill = "Genre",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          sort.val = "asc",# Sort the value in dscending order
          sort.by.groups = TRUE,     # Don't sort inside each group
          x.text.angle = 90,           # Rotate vertically x axis texts
          legend = "none",
          ggtheme = theme_pubclean())+
  font("x.text", size = 8, vjust = 0.5)+
  scale_x_discrete(breaks = seq(0,100,5))+labs(x = "2019", y = "")+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10))+
  theme(axis.text.x = element_blank(),  # Remove x-axis labels
        axis.ticks.x = element_blank())  # Remove x-axis tick marks  # Remove x-axis labels
#getting legend:
legend <- get_legend(y1)

#removing legend from y1
y1 <- y1 + theme(legend.position="none")+
  theme(axis.text.x = element_blank(),  # Remove x-axis labels
        axis.ticks.x = element_blank())  # Remove x-axis tick marks# Remove x-axis labels

#creating a blank
# 4. Create a blank plot
blankPlot <- ggplot()+geom_blank(aes(1,1)) + 
  cowplot::theme_nothing()

#Grid arranginging
grid.arrange(y1, y2,blankPlot, y3, y4,legend,
             ncol = 3,
             widths = c(2.7, 2.7,0.7), heights = c( 2.5, 2.5),
             top = textGrob(
               'Grossing in USD of 2016-19 Movies in different Genre',
               gp=gpar(fontsize=12,font=1, face = "bold"),
               hjust = 0.5,
               vjust = 1 )) # putting graphs together


ggplot(group_by(all.data,Genre), aes(x=Rank, y=Gross_earnings, fill=Genre)) +
  geom_bar(stat="identity", position=position_dodge())+
  geom_text(aes(label=NA), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)+
  scale_fill_brewer(palette="Paired")+
  labs( title = 'Grossing in USD of 2016-19 Movies in different Genre plotted against Rank',
        x = "Rank", 
        y = "Revenue (mn)")+
  theme(plot.title= element_text(size = 12),
        axis.text=element_text(size=8),
        axis.title=element_text(size=10))+
  theme_minimal()+
  facet_wrap(~Year)
```

```{r}

ggbarplot(all.data, x = "Rank", y = "Gross_earnings",
          fill = "Genre",               # change fill color by cyl
          color = "Genre",            # Set bar border colors to white
          palette = "aaas",            # "npg", "aaas", "lancet", "jco", "ucscgb", "uchicago", "simpsons" and                                                "rickandmorty" see ?ggpar
)+
  scale_x_discrete(breaks = seq(0, 100, 5))+
 labs(
    title = 'Movie Earnings by Title',
    subtitle = 'Movie Grossing in USD of 2016-19 Movies of different Genre plotted against Rank',
    y = "Revenue (mn)",
    x = "Rank"
  ) +
  font("x.text", size = 8, vjust = 0.5)+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10,face="bold"))
```

```{r}
Genres = c("Animation","Action","Drama","Comedy","Crime","Horror","Biography","Adventure")

ggbarplot( na.omit(all.data), x = "Genre", y = "Gross_earnings",
          fill = "Year",               # change fill color by cyl
          color = "Year",            # Set bar border colors to white
          palette = "uchicago",            # "npg", "aaas", "lancet", "jco", "ucscgb", "uchicago", "simpsons" and                                                "rickandmorty" see ?ggpar
)+
  scale_x_discrete()+
 labs(
    title = 'Movie Earnings by Genre',
    subtitle = 'Movie Grossing (USD) for Top 100 Popular Movies by Year and Genre'
  ) +
  ylab('Revenue (mn)')+
  font("x.text", size = 8, vjust = 0.5)+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10,face="bold"))
```

## 3 Examininging yearly trends
  3.1 Gross_revenue
```{r }
Genres = c("Animation","Action","Drama","Comedy","Crime","Horror","Biography","Adventure")

ggbarplot( na.omit(all.data), x = "Year", y = "Gross_earnings",
          fill = "Genre",               # change fill color by cyl
          color = "Genre",            # Set bar border colors to white
          palette = "aaas" # "npg", "aaas", "lancet", "jco", "ucscgb", "uchicago", "simpsons" and                                                "rickandmorty" see ?ggpar
)+
  scale_x_discrete()+
 labs(
    title = 'Movie Earnings by Year of Production',
    subtitle = 'Movie Grossing (USD) for Top 100 Popular Movies by Year and Genre'
  ) +
  ylab('Revenue (mn)')+
  font("x.text", size = 8, vjust = 0.5)+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10),
        plot.title = element_text(size = 12, vjust = 2, face = "bold"),
        plot.subtitle = element_text(size = 10, vjust = 2),
        legend.position = "bottom",
        legend.text = element_text(size = 8))

```

 3.2 Voting distribution
```{r}
ggplot(all.data, aes(x=Year, y=Votes/1000, fill=Genre)) +
  geom_boxplot()+
  theme_minimal()+
  labs(title = "Vote distribution across years")
```

3.3 Runtime distribution
```{r}
ggplot(all.data) +
  geom_boxplot(aes(x=Year, y=Runtime),fill = "violet", alpha = 0.3, col = "violet")+
  theme_minimal()+
  labs(title = "Film runtime distribution across years")+
  geom_boxplot(aes(x=Year, y=Runtime, fill=Genre))

```

3.4 Revenue distribution
```{r}
ggplot(all.data) +
  geom_boxplot(aes(x=Year, y=Gross_earnings),fill = "violet", alpha = 0.3, col = "violet")+
  theme_minimal()+
  labs(title = "Film Grossing distribution across years and genre", y = "Revenue (USD M)")+
  geom_boxplot(aes(x=Year, y=Gross_earnings, fill=Genre))

```

3.5 Revenue distribution
```{r }
Genres = c("Animation","Action","Drama","Comedy","Crime","Horror","Biography","Adventure")

ggplot(all.data, aes(x=Year, y=Gross_earnings)) +
  geom_violin(fill = "#7BCCB5",col = "#7BCCB5", alpha = 0.3 )+
  geom_violin(aes(fill = Genre))+
  geom_jitter(color = "#F79420", alpha = 0.3)+
  theme_minimal()+
  labs(title = "Movie Earnings distribution across years")
```

3.5 Runtime distribution
```{r }
Genres = c("Animation","Action","Drama","Comedy","Crime","Horror","Biography","Adventure")

ggplot(all.data, aes(x=Year, y=Runtime)) +
  geom_violin(fill = "#7BCCB5",col = "#7BCCB5", alpha = 0.3 )+
  theme_minimal()+
  labs(title = "Runtime distribution across years")

ggplot(all.data, aes(x=Year, y=Runtime)) +
  geom_violin(fill = "#7BCCB5",col = "#7BCCB5", alpha = 0.3 )+
  geom_jitter(aes(color = Genre, shape = Genre, width = 0.01, size = 0.1, alpha = 0.2))+
  theme_minimal()+
  labs(title = "Runtime distribution across years")+
  guides(size = FALSE, alpha = FALSE)

ggplot(all.data, aes(x=Year, y=Runtime)) +
  geom_violin(fill = "#7BCCB5",col = "#7BCCB5", alpha = 0.3 )+
  geom_violin(aes(fill = Genre))+
  geom_jitter(color = "#F79420", alpha = 0.3)+
  theme_minimal()+
  labs(title = "Runtime distribution across genre and years")
```


## 4. Exploring Top 10 Directors (by movie rating)
```{r}
directors <- all.data %>%
  group_by(Year) %>%
  arrange(Rank) %>%
  filter(row_number() <= 10) %>%
  ungroup()
directors

ggbarplot(subset(directors,Year==2016), x = "Rank", y = "Gross_earnings",
          fill = "Genre",               # change fill color by cyl
          color = "white",            # Set bar border colors to white
          palette = "jco",            # jco journal color palett. see ?ggpar
          #sort.val = "asc",# Sort the value in dscending order
          #sort.by.groups = TRUE,     # Don't sort inside each group
          legend = "right",
          x.text.angle = 90,          # Rotate vertically x axis texts,
          ggtheme = theme_pubclean())+
  font("x.text", size = 8, vjust = 0.5)+
  scale_x_discrete(breaks = seq(0,10,1))+labs(x = "Rank", y = "Revenue (mn)" )+
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=10))+
  geom_text(aes(label = Director, vjust = 0, angle = 90), size = 2.5)+
  labs(title = "Directors of 2016 Top 10 Movies by Rank")+
  theme(plot.title= element_text(size = 12, hjust = 0.5))
```
